---
title: "01-exercises"
author: "Christopher Brown"
date: "April 17, 2016"
output: html_document
---


## Exercise 1:

Write a function (f) that takes a vector of numbers, **x**, and returns a vector of numbers such that each element containing the product of every element of x except the element of x with the same index. 


### Example 

    > x <- c( 1, 5, 2, 8 ) 
    > f(x) 
    [1] 80 16 40 10
    # 5*2*8, 1*2*8, 1*5*8, 1*2*5  


### Solution 

```{r}

x <- c( 1, 5, 2, 8 ) 

prod(x) / x

f <- function() { 
  x <- c( 1, 5, 2, 8, 22 ) 
  
  for(i in x){
      print(i)
  }
}

f()

```
    
## Exercise 2

Write a function f(x) to accept an integer vector, and returns a vector with those numbers … except for: 
   multiples of 3 = “Fizz” 
   multiples of 5 = “Buzz”. 
   multiples of 3 and 5 = “FizzBuzz”
   
### Example

    > x <- 1:20
    > f(1:20) 
    # 1 2 Fizz 4 Buzz Fizz 7 8 Fizz Buzz 11 Fizz 13 14 FizzBuzz 16 17 Fizz 19 Buzz


### Solution

```{r}
#By Renaldo Williams

#Create vector of 20 numbers
x <- 1:20

#Create empty array vector
vector <- vector(mode="character", length=0)

#Create function to convert to Fizz, Buzz, or FizzBuzz
f <- function(numbers) {
  
  for (i in 1:length(numbers)) { 
    
    #Use modulus to find if multiple of 3
    if((numbers[i] %% 3 == 0) && (numbers[i] %% 5 == 0))
    {
      vector[i] <- "FizzBuzz"
    }
    #Use modulus to find if multiple of 5
    else if(numbers[i] %% 5 == 0)
    {
      vector[i] <- "Buzz"
    }
    #Use modulus to find if multiple of 3 and 5
    else if(numbers[i] %% 3 == 0)
    {
      vector[i] <- "Fizz"
    }
    else {
      vector[i] <- numbers[i]
    }
  }
  
  vector
}

# run the function
f(x)
```
