
title: "05-exercises"
author: "Your Name Here"
date: "2016-05-xx"
output: html_document
---
  
  ## Reading:
  - **APM** Chapter 8.1-8.5 "Regression Trees and Rule-Based Models" (25 pages)
- **APM** Chapter 14.1-14.5 "Classification Trees and Rule-Based"  

```{r, echo=FALSE, results='hide', warning=FALSE }
packs <-  c('ggplot2', 'magrittr', 'dplyr', 'caret', 'AppliedPredictiveModeling')

for( nm in packs ) { 
  # message(nm)
  if( ! nm  %in% installed.packages()[,1]  ) install.packages(nm)
  library(nm, character.only = TRUE)
}

.. = NULL  # For Aesthetics

```


## Exercise 1: GermanCredit

Revisit the GermanCredit data. Use `caret` to build models of `Class` using the following techniques:
  
  - glm
- rpart
- knn
- party::ctree
- randomForest
- A method of your choice from the Caret Model List (you will need to install any dependencies)

Save the   caret objects with the names provided.

```{r}

data("GermanCredit")

# (RW) First thing, ALWAYS, ALWAYS spend 5 minutes analyzing CVS file in excel to gain familiarity with data features and labels before building classifiers.

# Create a CSV GermanCredit file
#x <- data.frame(GermanCredit)
#write.csv(x, file = "GermanCredit.csv")


# Your work here. 

# Have to convert class to factor for regression model since models need number not text.
data("GermanCredit")
GermanCredit$Class <- factor(GermanCredit$Class)
#GermanCredit$Class <- as.numeric(GermanCredit$Class)

# use sample() to select 1000 random numbers
random_sample <- sample(1000, 1000)
train_indexes <- random_sample[1:700]
test_indexes <- random_sample[701:1000]

length(train_indexes)
length(test_indexes)

germancredit_train <- GermanCredit[train_indexes,]
germancredit_test <- GermanCredit[test_indexes,]

germancredit_train_labels <- germancredit_train$Class
germancredit_test_labels <- germancredit_test$Class

prop.table(table(GermanCredit$Class))
prop.table(table(germancredit_train$Class))
prop.table(table(germancredit_test$Class))

str(germancredit_train)

# Using caret and train() for automatic parameter tuning. Train serves as standardized interface for over 175 different machine learning models

tc <- trainControl( method="boot", number=5, classProb=TRUE, savePrediction=TRUE )

#get column index of class
classColumnIndex <- grep("Class", colnames(germancredit_train))

# Getting 12 warnings: prediction from a rank-deficient fit may be misleading
fit.glm <- train(germancredit_train$Class ~ ., data = germancredit_train[,-classColumnIndex], method = "glm", family = "binomial")

str(germancredit_train)

germancredit_train %>% View

# Trying with regular glm, Train the regression model with training data
fit.glm <- glm(germancredit_train$Class ~ ., data = germancredit_train, family = "binomial")

# Now predict using test data
# Getting error: prediction from a rank-deficient fit may be misleading
fit.pred <- predict(fit.glm, newdata = germancredit_test)

summary(fit.glm)

# Attempting Classification with KNN 
# Since using KNN, need to normalize features since distance calculation of KNN is heavily dependent upon the measurement scale of input features

normalize <- function(x) {
  return ( (x - min(x)) / (max(x) - min(x)) )
}

#test it
normalize(c(1,2,3,4, 5))
gc <- GermanCredit

#convert category to factor
gc$Class <- factor(gc$Class)

#now lets normalize

#get column index of class
classColumn <- grep("Class", colnames(gc))

#normalize every column but Class column
germancredit_n <- as.data.frame(lapply(gc[,-classColumn], normalize)) 

#add back Class column
germancredit_n$Class <- gc[,classColumn]

str(germancredit_n)

#Remove NaN
#Retrieve list of NaN columns
nan <- sapply(germancredit_n, function(x) all(is.nan(x)))
#Remove NaN columns
germancredit_n <- germancredit_n[, !nan]

#create training and test set
germancredit_n_train <- germancredit_n[train_indexes,]
germancredit_n_test <- germancredit_n[test_indexes,]

prop.table(table(germancredit_n$Class))
prop.table(table(germancredit_n_train$Class))
prop.table(table(germancredit_n_test$Class))

#Get labels for training and test set
germancredit_n_train_labels <- germancredit_n_train$Class
germancredit_n_test_labels <- germancredit_n_test$Class

library("class")

#Get Class columns for training and test set
trCol <- grep("Class", colnames(germancredit_n_train))
testCol <- grep("Class", colnames(germancredit_n_test))

#Run KNN classifier with labels removed from training and test data
germancredit_test_pred <- knn(train=germancredit_n_train[,-trCol], test=germancredit_n_test[,-testCol], cl=germancredit_n_train_labels, k=21)
?knn
summary(germancredit_test_pred)

library("gmodels")
CrossTable(x = germancredit_n_test_labels, y=germancredit_test_pred, prop.chisq=FALSE)

confusionMatrix(germancredit_test_pred, germancredit_n_test_labels, positive="Bad")

#Error rate of ~29% so performance is not good. Lets investigate ROC

#install.packages("ROCR")
library("ROCR")

###TRAIN KNN###
germancredit_test_pred_prob <- knn(train=germancredit_n_train[,-trCol], test=germancredit_n_test[,-testCol], cl=germancredit_n_train_labels, k=21, prob=TRUE)

pred_prob_values <- attr(germancredit_test_pred_prob,"prob")

pred <- prediction(predictions = pred_prob_values, labels = germancredit_n_test_labels)

perf <- performance(pred, measure="tpr", x.measure="fpr")
plot(perf,main="ROC Curve for GermanCredit data", col="blue", lwd=3)
abline(a=0,b=1,lwd=2,lty=2)

perf.auc <- performance(pred, measure="auc")
unlist(perf.auc@y.values)

str(germancredit_n_train)
str(germancredit_n_test)
?knn

germancredit_train
germancredit_test
#remove class before normalize
germancredit_train$Class <- NULL
#normalize
germancredit_train_n <- as.data.frame(lapply(germancredit_train, normalize)) 


table(germancredit_test$Personal.Female.Single)

fit.knn <- 
  
  fit.knn <- train(germancredit_n_train$Class ~ ., data = germancredit_n_train[,-trCol], method="knn")
str(fit.knn)

fit.knnpred <- predict(fit.knn, germancredit_n_test[,-testCol])

fit.knnpred_prob <- predict(fit.knn, germancredit_n_test[,-testCol], type="prob")
head(fit.knnpred_prob)
confusionMatrix(fit.knnpred, germancredit_n_test_labels, positive="Bad")



fit.rpart <- ..
fit.rf <- ..
fit.myown <- ..


```


- Compare the models using `caret::confusionMatrix`
- Comparing the models Using the `pROC` packages
- create ROC curves for the models 

Show your work! 
  
  ```{r}

.. # YOUR WORK HERE

```


Q: Which models would you select based on these tools?

Q: If you assume that a `Class=="bad""` is 10 more costly than `Class=="good"`, determine your threshold for the model of your choice.  Show your work.


```{r}

.. # YOUR WORK HERE


```
