---
title: "02-exercises"
instructor: Christopher Brown
student: Renaldo Williams
date: "April 13, 2016"
output: pdf_document
---

This exercise uses the **Fuel Economy** data set from the **AppliedPredicitiveModeling** package.

Note: The following will set-up your environment for this exercise. If you get an error stating that the packages have not been found, you need to install those packages.


```{r,echo=FALSE}

packs <-  c('AppliedPredictiveModeling', 'ggplot2', 'magrittr', 'dplyr')

for( nm in packs ) { 
  # message(nm)
  if( ! nm  %in% installed.packages()[,1]  ) install.packages(nm)
  library(nm, character.only = TRUE)
}

# Load data set into environment
data(FuelEconomy)
... = NULL  # Needed for aesthetics 

```

## Exercise 1

Hint: See `?cars2010`


- After the **Fuel Economy** data is loaded, combine three data sets into one data set. (Note: The name `dat` is very often used in these situations, `data` is a reserved R word.)  

```{r}
dat3years <- rbind(cars2010,cars2011,cars2012)

naive_guess = mean(dat3years$FE)

set.seed(314)

samp <- dat3years %>% dplyr::sample_n(10) # you can also do just sample_n(10) since no conflicts with deep lyer package

samp %>% ggplot(aes(x=EngDispl, y=FE) ) + geom_point()

#Naive guess (red line) is just a good place to start, always need a good place to start
samp %>% ggplot(aes(x=EngDispl, y=FE) ) + geom_point() + geom_hline(yintercept=naive_guess, color="red")

#graph showing naive guess and linear model
samp %>% ggplot(aes(x=EngDispl, y=FE) ) + geom_point() + geom_hline(yintercept=naive_guess, color="red") + stat_smooth(method = "lm", se=FALSE)

plot(samp$FE - naive_guess)
qplot(samp$FE - naive_guess)

#dat3years$FE

(dat3years$FE - naive_guess)^2 %>% mean %>% sqrt

fit.dat3years <- lm(FE ~ EngDispl, data = dat3years)

fit.dat3years

fit.dat3years <- lm(FE ~ EngDispl + NumCyl, data = dat3years)

fit.dat3years

fit.dat3years <- lm(FE ~ EngDispl + NumCyl + NumGears, data = dat3years)

fit.dat3years

fit.dat3years <- lm(FE ~ EngDispl + CarlineClassDesc + EngDispl, data = dat3years)

fit.dat3years

fit.dat3years %>% summary()


#dot means use all variables 
fit.dat3years <- lm(FE ~ . + EngDispl, data = dat3years)

fit.dat3years

#sample 10 CarlineClassDesc
dat3years %>% select(CarlineClassDesc) %>% sample_n(10)

#print table count for each CarlineClassDesc type
dat3years %>% select(CarlineClassDesc) %>% table

```


- What is a good "naive guess" of FE? Show your work

```{r}

dat3years <- rbind(cars2010,cars2011,cars2012)

naive_guess = mean(dat3years$FE)

naive_guess

```


- plot FE (Fuel Econonomy) vs. EngDisp. Plot the naive guess.

```{r}
# ... ggplot2  

dat3years %>% ggplot(aes(x=EngDispl, y=FE) ) + geom_point()

```

- Sample 10 observations from `dat`
- Plot this data. Add a line for the naive_guess. 


```{r}

set.seed(314) 

# Sample 

samp <- dat3years %>% dplyr::sample_n(10) 

#Naive guess (red line) is just a good place to start, always need a good place to start
samp %>% ggplot(aes(x=EngDispl, y=FE) ) + geom_point() + geom_hline(yintercept=naive_guess, color="red")


```


## Exercise 2:


Write a loss functions for calculating:

- Root Mean Square Error
- Mean Absolute Error
- Median Absolute Error

All functions should accept two arguments: 

```{r}

rmse <- function(y,yhat) {

  ( y - yhat )^2  %>% mean %>% sqrt 
}

mae <- function(y, yhat) {
  
  abs( y - yhat ) %>% mean()
}

medae <- function(y, yhat) { 
  
  abs( y - yhat ) %>% median()
}

```

Use these functions to evaluate the loss/performance of:
- the naive guess

## Exercise 3: Linear Model and Model Performance 

- Use `lm` to create a linear model fitting the relationship between FE and EngDispl for the cars2010 data set


```{r}

fit.2010 <- lm( FE ~ EngDispl, data=cars2010 )

```

- Use your functions to evaluate the training error
- Use your model to:
-- predict the FE for 2011. What is the RMSE errors associated with the predictions.
-- predict the FE for 2012. What is the RMSE errors associated with the predictions.

```{r}

#Predict FE for 2010, 2011, 2012 using lm from 2010
y.2010 <- predict( fit.2010, data=cars2010 )
y.2011 <- predict( fit.2010, data=cars2011 )
y.2012 <- predict( fit.2010, data=cars2012 )

#Calculate RMSE error
rmse.2010 <- rmse( cars2010$FE,y.2010)
rmse.2011 <- rmse( cars2011$FE,y.2011)
rmse.2012 <- rmse( cars2012$FE,y.2012)

```


```{r}
# DO NOT EDIT 
rmse.2010 
rmse.2011
rmse.2012
```


## Exercise 4: 

- Model the fuel economy (`FE`) as a function of `EngDispl`, `NumCyl` and `VarValve` using the `cars2011` data set.  
- Provide betas


```{r}
fit.2011 <- lm( FE ~ EngDispl + NumCyl + VarValveTiming, data=cars2011 ) 

summary(fit.2011)

coef(fit.2011)

```
